---
title: "R Notebook"
output: html_notebook
---
```{r}
library(data.table)
library(lubridate)
library(tidyverse)
library(stargazer)
library(tidymodels)
library(rpart.plot)
library(shiny)
library(ADNIMERGE)
library(plyr)
library(writexl)
library(zoo)
```


```{r}
adni.new <- read.csv("/Users/jinchenxie/Desktop/ML_Pipeline/project/df_ohe_corrected.csv")
# adni.new <- read.csv("/Users/jinchenxie/Desktop/df_dx_imputed.csv")

```

```{r}
adni.ids <- unique(adni.new$RID)

adni.ids.2 <- unique(adni.data$RID)

new.ids <- adni.ids[which(!(adni.ids %in% adni.ids.2))]

adni.new.points <- adni.new[which(adni.new$RID %in% new.ids),]
adni.newpoints.first <- ddply(adni.new.points,.(RID), head,1)


write.csv(adni.newpoints.first,'/Users/jinchenxie/Desktop/ML_Pipeline/project/new_points.csv')
```


```{r}
table(adni.new$DX)
table(adni.new$DX_bl)
```

```{r}
# adni.new <- adni.new %>% mutate(DX_bl_2= ifelse((DX_bl=="SMC"|DX_bl=="CN"), 0, 1))
# adni.new <- adni.new %>% mutate(DX_bl_2= ifelse(DX_bl=="AD", 2, DX_bl_2))
# 
# 
# adni.new <- adni.new %>% mutate(DX_2= ifelse(DX=="CN", 0, 1))
# adni.new <- adni.new %>% mutate(DX_2= ifelse(DX=="Dementia", 2, DX_2))
# 
# adni.new <- adni.new %>% mutate(DX_bl=DX_bl_2)
# adni.new <- adni.new %>% mutate(DX=DX_2)
```


0: CN
1: MCI
2: AD


# Include at least 24 months patients only
```{r}
rows.24 = (adni.new[which(adni.new$Month >= 24),])
length(unique(rows.24$RID))
rid.atleast.24 = unique(rows.24$RID)
adni.24 = adni.new[which(adni.new$RID %in% rid.atleast.24),]
```


There are 863 participants who have at least 24 months follow-ups.  
`adni.24` contains all records for participants who had at least 24 months follow-ups.


```{r save dataset version 1}
write_xlsx(adni.24, "/Users/jinchenxie/Desktop/ML_Pipeline/project/adni_24_v1.xlsx")
```


## participants have 6, 12, 24-month label
```{r}
dim(adni.24[which(adni.24$Month == 3),])[1]
dim(adni.24[which(adni.24$Month == 6),])[1]
dim(adni.24[which(adni.24$Month == 12),])[1]
dim(adni.24[which(adni.24$Month == 24),])[1]
dim(adni.24[which(adni.24$Month == 36),])[1]
```

There are 592 participants who had at least 24 months follow-ups and had DX label at the 3-month time-point.  
There are 658 participants who had at least 24 months follow-ups and had DX label at the 6-month time-point.  
There are 766 participants who had at least 24 months follow-ups and had DX label at the 12-month time-point.  
There are 798 participants who had at least 24 months follow-ups and had DX label at the 24-month time-point.  
There are 347 participants who had at least 24 months follow-ups and had DX label at the 36-month time-point.  

```{r}
rid.3 = unique(adni.new[which(adni.new$Month == 3),]$RID)
rid.6 = unique(adni.new[which(adni.new$Month == 6),]$RID)
rid.12 = unique(adni.new[which(adni.new$Month == 12),]$RID)
rid.24 = unique(adni.new[which(adni.new$Month == 24),]$RID)
rid.36 = unique(adni.new[which(adni.new$Month == 36),]$RID)
```

```{r}
length(unique(adni.new[which((adni.new$RID %in% rid.3)&(adni.new$RID %in% rid.6)&(adni.new$RID %in% rid.12)),]$RID))

length(unique(adni.new[which((adni.new$RID %in% rid.6)&(adni.new$RID %in% rid.12)),]$RID))

length(unique(adni.new[which((adni.new$RID %in% rid.12)&(adni.new$RID %in% rid.24)),]$RID))

length(unique(adni.new[which((adni.new$RID %in% rid.6)&(adni.new$RID %in% rid.24)),]$RID))
length(unique(adni.new[which((adni.new$RID %in% rid.6)&(adni.new$RID %in% rid.12)&(adni.new$RID %in% rid.24)),]$RID))
length(unique(adni.new[which((adni.new$RID %in% rid.12)&(adni.new$RID %in% rid.24)&(adni.new$RID %in% rid.36)),]$RID))
length(unique(adni.new[which((adni.new$RID %in% rid.3)&(adni.new$RID %in% rid.6)&(adni.new$RID %in% rid.12)&(adni.new$RID %in% rid.24)),]$RID))
```
635 participants had 3-month, 6-month and 12-month labels.  
1425 participants had both 6-month and 12-month labels. 
1238 participants had both 6-month and 24-month labels.  
1326 participants had both 12-month and 24-month labels.  
1166 participants had 6-month, 12-month and 24-month labels. 
730 participants had both 12-month, 24-month and 36-month labels.  
505 participants had 3-month, 6-month, 12-month and 24-month labels.  

Use only participants with 6-month, 12-month and 24-month labels.
```{r}
rid.3.6.12 = unique(adni.new[which((adni.new$RID %in% rid.6)&(adni.new$RID %in% rid.12)&(adni.new$RID %in% rid.3)),]$RID)
adni.3.6.12 = adni.new[which(adni.new$RID %in% rid.3.6.12),]

rid.6.12.24 = unique(adni.new[which((adni.new$RID %in% rid.6)&(adni.new$RID %in% rid.12)&(adni.new$RID %in% rid.24)),]$RID)
adni.6.12.24 = adni.new[which(adni.new$RID %in% rid.6.12.24),]

rid.12.24.36 = unique(adni.new[which((adni.new$RID %in% rid.12)&(adni.new$RID %in% rid.24)&(adni.new$RID %in% rid.36)),]$RID)
adni.12.24.36 = adni.new[which(adni.new$RID %in% rid.12.24.36),]
```

```{r}
# adni.24.mci <- adni.24[which(adni.24$DX_bl==1),]
adni.6.12.24.mci <- adni.6.12.24[which(adni.6.12.24$DX_bl==1),]
adni.6.12.24.cn <- adni.6.12.24[which(adni.6.12.24$DX_bl==0),]

adni.12.24.36.mci <- adni.12.24.36[which(adni.12.24.36$DX_bl==1),]

adni.3.6.12.mci <- adni.3.6.12[which(adni.3.6.12$DX_bl==1),]
length(unique(adni.12.24.36.mci$RID))
```

`adni.6.12.24`: participants had non-missing DX at 6, 12, 24 months visit.  


## what if we only consider patients started with CN or MCI?

```{r}
adni.first <- ddply(adni.new,.(RID), head,1)
table(adni.first$DX_bl)
```

There are 450 participants started with CN at their baseline visits.  
There are 544 participants started with MCI at their baseline visits.  

```{r}
adni.cn.rid <- adni.first[which(adni.first$DX_bl==0),]$RID
adni.mci.rid <- adni.first[which(adni.first$DX_bl==1),]$RID
adni.cn = adni.new[which(adni.new$RID %in% adni.cn.rid),]
adni.mci = adni.new[which(adni.new$RID %in% adni.mci.rid),]
```


Among participants started with CN, how many (%) had 3, 6, 12, 24-month check-point labels? 
```{r}
length(unique(adni.cn[which(adni.cn$Month == 3),]$RID))
length(unique(adni.cn[which(adni.cn$Month == 3),]$RID))/length(unique(adni.cn$RID))
length(unique(adni.cn[which(adni.cn$Month == 6),]$RID))
length(unique(adni.cn[which(adni.cn$Month == 6),]$RID))/length(unique(adni.cn$RID))
length(unique(adni.cn[which(adni.cn$Month == 12),]$RID))
length(unique(adni.cn[which(adni.cn$Month == 12),]$RID))/length(unique(adni.cn$RID))
length(unique(adni.cn[which(adni.cn$Month == 24),]$RID))
length(unique(adni.cn[which(adni.cn$Month == 24),]$RID))/length(unique(adni.cn$RID))
```

Among participants started with MCI, how many (%) had 3, 6, 12, 24-month check-point labels? 
```{r}
length(unique(adni.mci[which(adni.mci$Month == 3),]$RID))
length(unique(adni.mci[which(adni.mci$Month == 3),]$RID))/length(unique(adni.mci$RID))
length(unique(adni.mci[which(adni.mci$Month == 6),]$RID))
length(unique(adni.mci[which(adni.mci$Month == 6),]$RID))/length(unique(adni.mci$RID))
length(unique(adni.mci[which(adni.mci$Month == 12),]$RID))
length(unique(adni.mci[which(adni.mci$Month == 12),]$RID))/length(unique(adni.mci$RID))
length(unique(adni.mci[which(adni.mci$Month == 24),]$RID))
length(unique(adni.mci[which(adni.mci$Month == 24),]$RID))/length(unique(adni.mci$RID))
```







# Add point and transition label (6, 12, 24 point labels and transitions)
## 0, 3, 6, 12, 24, 36 point label
```{r}
adni.point.label <- adni.6.12.24.mci %>% mutate(m0_label=DX_bl)

adni.point.label <- adni.point.label %>% group_by(RID) %>%  mutate(m3_label=ifelse(Month==3, DX, NA)) %>% 
  group_by(RID) %>% 
  fill(m3_label) %>% #default direction down
  fill(m3_label, .direction = "up") 

adni.point.label <- adni.point.label %>% group_by(RID) %>%  mutate(m6_label=ifelse(Month==6, DX, NA)) %>% 
  group_by(RID) %>% 
  fill(m6_label) %>% #default direction down
  fill(m6_label, .direction = "up")

adni.point.label <- adni.point.label %>% group_by(RID) %>%  mutate(m12_label=ifelse(Month==12, DX, NA)) %>% 
  group_by(RID) %>% 
  fill(m12_label) %>% #default direction down
  fill(m12_label, .direction = "up")

adni.point.label <- adni.point.label %>% group_by(RID) %>%  mutate(m18_label=ifelse(Month==18, DX, NA)) %>% 
  group_by(RID) %>% 
  fill(m18_label) %>% #default direction down
  fill(m18_label, .direction = "up")

adni.point.label <- adni.point.label %>% group_by(RID) %>%  mutate(m24_label=ifelse(Month==24, DX, NA)) %>% 
  group_by(RID) %>% 
  fill(m24_label) %>% #default direction down
  fill(m24_label, .direction = "up")

adni.point.label <- adni.point.label %>% group_by(RID) %>%  mutate(m30_label=ifelse(Month==30, DX, NA)) %>% 
  group_by(RID) %>% 
  fill(m30_label) %>% #default direction down
  fill(m30_label, .direction = "up")

adni.point.label <- adni.point.label %>% group_by(RID) %>%  mutate(m36_label=ifelse(Month==36, DX, NA)) %>% 
  group_by(RID) %>% 
  fill(m36_label) %>% #default direction down
  fill(m36_label, .direction = "up")

adni.point.label <- adni.point.label %>% group_by(RID) %>%  mutate(m48_label=ifelse(Month==48, DX, NA)) %>% 
  group_by(RID) %>% 
  fill(m48_label) %>% #default direction down
  fill(m48_label, .direction = "up")
```

## transition label
Transition from 0 -> 1
```{r}
adni.label <- adni.point.label %>% group_by(RID) %>% mutate(tran_0to1_m3=ifelse((m0_label==0 & m3_label==1), 1, 0))
adni.label <- adni.label %>% group_by(RID) %>% mutate(tran_0to1_m6=ifelse((m0_label==0 & m6_label==1), 1, 0))
adni.label <- adni.label %>% group_by(RID) %>% mutate(tran_0to1_m12=ifelse((m0_label==0 & m12_label==1), 1, 0))
adni.label <- adni.label %>% group_by(RID) %>% mutate(tran_0to1_m24=ifelse((m0_label==0 & m24_label==1), 1, 0))
adni.label <- adni.label %>% group_by(RID) %>% mutate(tran_0to1_m36=ifelse((m0_label==0 & m36_label==1), 1, 0))
```

Transition from 1 -> 2
```{r}
adni.label <- adni.point.label %>% group_by(RID) %>% mutate(tran_1to2_m3=ifelse((m0_label==1 & m3_label==2), 1, 0))
adni.label <- adni.label %>% group_by(RID) %>% mutate(tran_1to2_m6=ifelse((m0_label==1 & m6_label==2), 1, 0))
adni.label <- adni.label %>% group_by(RID) %>% mutate(tran_1to2_m12=ifelse((m0_label==1 & m12_label==2), 1, 0))
adni.label <- adni.label %>% group_by(RID) %>% mutate(tran_1to2_m18=ifelse((m0_label==1 & m18_label==2), 1, 0))
adni.label <- adni.label %>% group_by(RID) %>% mutate(tran_1to2_m24=ifelse((m0_label==1 & m24_label==2), 1, 0))
adni.label <- adni.label %>% group_by(RID) %>% mutate(tran_1to2_m30=ifelse((m0_label==1 & m30_label==2), 1, 0))
adni.label <- adni.label %>% group_by(RID) %>% mutate(tran_1to2_m36=ifelse((m0_label==1 & m36_label==2), 1, 0))
adni.label <- adni.label %>% group_by(RID) %>% mutate(tran_1to2_m48=ifelse((m0_label==1 & m48_label==2), 1, 0))
```

```{r}
adni.label <- adni.label %>% mutate(tran_1to2_m6to12=ifelse((tran_1to2_m12==1 & tran_1to2_m6==0), 1, 0))
adni.label <- adni.label %>% mutate(tran_1to2_m12to24=ifelse((tran_1to2_m24==1 & tran_1to2_m12==0), 1, 0))
```

Transition from 0 -> 2
```{r}
adni.label <- adni.point.label %>% group_by(RID) %>% mutate(tran_0to2_m3=ifelse((m0_label==0 & m3_label==2), 1, 0))
adni.label <- adni.label %>% group_by(RID) %>% mutate(tran_0to2_m6=ifelse((m0_label==0 & m6_label==2), 1, 0))
adni.label <- adni.label %>% group_by(RID) %>% mutate(tran_0to2_m12=ifelse((m0_label==0 & m12_label==2), 1, 0))
adni.label <- adni.label %>% group_by(RID) %>% mutate(tran_0to2_m24=ifelse((m0_label==0 & m24_label==2), 1, 0))
adni.label <- adni.label %>% group_by(RID) %>% mutate(tran_0to2_m36=ifelse((m0_label==0 & m36_label==2), 1, 0))
```

```{r}
adni.label.first <- ddply(adni.label,.(RID), head,1)
adni.label.first %>% select(starts_with("tran_")) %>% mutate_all(as.factor) %>% summary()
```


```{r}
write_xlsx(adni.label, "/Users/jinchenxie/Desktop/ML_Pipeline/project/adni_label_v3.xlsx")
write_xlsx(adni.label.first, "/Users/jinchenxie/Desktop/ML_Pipeline/project/adni_label_v5.xlsx")
write.csv(adni.label.first,'/Users/jinchenxie/Desktop/ML_Pipeline/project/adni_label_v5.csv')
```

```{r}
adni.label %>% select(RID, Month, DX_bl, DX, starts_with("tran_")) %>% filter(is.na(tran_1to2_m24))
```

# Last two weeks - label processing

```{r}
adni.label <- read.csv("/Users/jinchenxie/Desktop/ML_Pipeline/project/adni_label_v6.csv")
```

```{r}
adni.label <- adni.label %>% mutate(m6_ind=ifelse(tran_1to2_m6==1, 0, 1))
adni.label <- adni.label %>% mutate(m12_ind=ifelse((tran_1to2_m6==1 | tran_1to2_m6to12==1), 0, 1))
```

```{r}
model.label.6 <- adni.label %>% filter(m6_ind==1)
model.label.12 <- adni.label %>% filter(m12_ind==1)

```

```{r}
write.csv(adni.label,'/Users/jinchenxie/Desktop/ML_Pipeline/project/adni_label_wind.csv')
write.csv(model.label.6,'/Users/jinchenxie/Desktop/ML_Pipeline/project/adni_label_6m_early.csv')
write.csv(model.label.12,'/Users/jinchenxie/Desktop/ML_Pipeline/project/adni_label_12m_early.csv')
```

**Check label % consistency at each time point**
(e.g. could it be only patients with AD will get to be tested again at 6-month time point?)

```{r}
table(adni.new$Month)
```

```{r}
adni.label <- read.csv("/Users/jinchenxie/Desktop/ML_Pipeline/project/adni_label_wind.csv")
```

```{r}
adni.label %>% filter(m12_ind==1) 
```
